River networks dampen long-term hydrological signals of climate change

### 2017-06-02 drafts/AGU_Journal/submission 2/ms_Geo_Let.tex
Removed 75+ words from MS to meet GPRL word limit and added a highlighted version of the MS to the .gitignore file.

%Packages
\usepackage{upgreek}
\usepackage{gensymb}

@@

\begin{abstract}
	River networks may dampen local hydrologic climate change through the aggregation of upstream climate portfolio assets. Here we examine this hypothesis using flow and climate trend estimates (1970--2007) at 55 hydrometric gauge stations and across their contributing watersheds' within the Fraser River basin in British Columbia, Canada. Using a null hypothesis framework we compared our observed attenuation of river flow trends as a function of increasing area and climate trend diversity, with null-simulated estimates to gauge the likelihood and strength of our observations. We found the Fraser River reduced variability in downstream long-term discharge by $>$91\%, with $>$3.1 times the attenuation than would be expected under null-simulation. Although the strength of dampening varied seasonally, our findings indicate that large free-flowing rivers offer a powerful and largely unappreciated process of climate change mitigation. River networks that integrate a diverse climate portfolio can dampen local extremes and offer climate change relief to riverine biota.
\end{abstract}

@@

\begin{figure}[h]
\centering
\includegraphics[width=35pc]{Fig1_Map.pdf}
	\caption{Climate trends (1970 - 2007) in mean annual temperature (MAT) and mean annual precipitation (MAP) within the Fraser River basin overlayed on a digital elevation model of British Columbia, Canada. Flow gauge sites (dots) are scaled by the size of the contributing area. The map is projected in Albers BC UTMs to provide an equal area depiction of the region but labels are expressed in WGS84 latitude and longitude.}
\label{fig:1}
\end{figure}

Here we consider whether river networks attenuate local long-term climate change through analysis of hydrometric data from one of the largest free-flowing rivers in North America. Located in British Columbia, Canada, the Fraser River drains an area approximately the size of the United Kingdom ($\sim$217,000 km\textsuperscript{2}) and in a rare combination is fairly well-monitored while also having no dams on its main stem \citep{Vorosmarty:2010} (Figure~\ref{fig:1}). This watershed drains interior high plateau, coastal mountains, and the Canadian Rockies, and thus integrates a diverse mosaic of landscapes, weather, and climate. As with other mid-latitude rivers \citep{Bindoff:2013}, the Fraser River's discharge volume and variability have increased over the last several decades \citep{Dery:2012,Morrison:2002}. Following projected pattens of climate change, the Fraser River basin is increasingly shifting from a snow dominated to rain-snow hybrid system \citep{Kang:2016}. Using 38 years of flow data collected concurrently at 55 hydrometric gauge stations throughout the watershed, we use a novel analytical approach to examine whether this large river exhibits signals of long-term climate dampening.

@@

	To assess signals of climate change in hydrology within the Fraser River basin we downloaded daily flow data from Environment Canada's HYDAT database. Within HYDAT we selected Fraser River basin hydro-metric gauge stations that were not observed to be dam influenced and which collected data in each month between 1970 and 2007; where no month was missing more than 5 days data and at least 35 complete years were present. We selected this 38-year range because this period maximized the number of gauge stations in the Fraser River basin (\textit{n}=55) operating concurrently over a long enough time period in which climate trends may be observable. We linearly interpolated missing data in log-space, but of the 746,544 days of data only 23 days were interpolated. These daily data were smoothed using a five-day rolling average before summarizing into annual and monthly response variables. This smoothing technique was used to reduce the influence of erroneous and unusually extreme values. Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and sites that were observed downstream of, and atypically close to, a dam (\textit{n}=3). The downstream distance of gauge stations from dams was determined using GIS spatial layers available in the \textit{BC Data Catalog}, maintained by the provincial government of British Columbia.

	As climate change shifts precipitation patterns from snow to rain in temperate regions, rivers are generally predicted to have lower low-flows and higher high-flows, as well as earlier snowmelt \citep{Nijssen:2001}. These manifestations of rising temperatures are impacting species \citep{Xenopoulos:2006} and people \citep{Hirabayashi:2013}. We summarized our daily rolling average discharge data into annual and monthly flow metrics that capture these transitions. We used maximum- and minimum-flow estimates to capture trends in extreme discharge and median-flow estimates to capture general annual and seasonal trends. Because the timing of flow is particularly important to the phenology of many organisms, we also calculated the day-of-year in which half the annual flow was reached.

@@

	Spatial variation in climate change would suggest catchments of different size integrate different amounts of climate variability. In order to calculate a climate index for each gauge station we delineated flow-gauge catchments and summarized climate trends within those catchments. Using Whitebox Geospatial Analysis Tools \citep{Lindsay:2016}, we delineated the Fraser River basin and the catchment areas of each flow-gauge station. Digital elevation models were provided by the provincial government of British Columbia, Canada and state government of Washington, USA and were pre-processed with a breaching algorithm and stream burned to facilitate proper flow path and accumulation models \citep{Woodrow:2016}. We summarized the climate variability integrated at each flow gauge by calculating the standard deviation in climate trends within their catchments. The standard deviations of these climate variables were scaled between 0 and 1 and summed as a general climate index for each catchment.

@@

We hypothesize that flow gauge sites with larger climate portfolios will exhibit greater long-term flow trend stability than sites with smaller climate portfolios. We used catchment area as a proxy for the climate portfolio diversity of a given site, given that larger catchments have more variable climate portfolios (Figure~\ref{fig:2}) as well as aggregate more water and carry more weight on downstream dampening. Therefore, to test whether networks dampen climate signals, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
\begin{linenomath*}
\begin{equation}
	\hat{b}_{s} = c + d\sqrt{\mathrm{A}_{s}} + \eta_{s}, \quad
  \eta_{s} \sim \mathcal{N}(0, f(\mathrm{A}_{s})) \label{eq3},
\end{equation}
\end{linenomath*}
where $\hat{b}_{s}$ represents a flow trend at a given site ($s$) and $\mathrm{A}_{s}$ represents site $s$'s watershed area. We adjusted $\sqrt{\mathrm{A}_{s}}$ by subtracting the mean $\sqrt{\mathrm{A}}$, thereby centering our predictor. Fitted $d$ and $c$ parameters represent the mean effect of watershed area ($\sqrt{\mathrm{A}}$), on flow trends and the mean flow trend for a $\sim$5,700 km\textsuperscript{2} watershed, respectively. Captured in our climate portfolio proxy variable ($\mathrm{A}$), is the interaction between climate and landuse. Undoubtedly, timber harvest, the dominant source of land cover change in the Fraser River basin, scales with sub-basin area and contributes to changes in flow. Although exploratory linear models at the site level do suggest an effect of timber harvest on long-term flow trends for some gauge stations (Figure~S1, S2), when information is shared across sites in a global model, climate effects dominate (Figure~S3). Thus we did not attempt to separate the contributions of land cover change, climate change and their interaction on long-term flow trends.

@@

Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 92 and 96\% dampening of flow trends moving from headwater catchments to the network's confluence. In other words, the largest catchments were approximately 10 times more stable in their climate response than the smallest catchments. For instance, trends ranged between a 6\% reduction and 19\% increase in median flow per decade among small watersheds, while large watersheds ranged between 0.8 and 1.6\% change per decade. Similar attenuation was seen in flow timing with a 94\% reduction in the day-of-year to half-annual-flow trends such that small watersheds were reaching half their annual flow between 8 days later and 25 days earlier while large watersheds have only shifted 2 to 4 days earlier between 1970 and 2007.

\begin{figure}[h]
\centering
\includegraphics[width=35pc]{Fig4_AnnFun.pdf}
	\caption{Annual flow trend attenuation within the Fraser River basin. (\textbf{Left}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large). Blue lines represent observed attenuation; orange and red represent simulated attenuation that is weaker and stronger than observed. (\textbf{Right}) Density plots show null model simulated variance exponents ($\hat{\updelta}$) and the proportion on either side of the observed variance exponent (blue). Flow metrics include long-term flow-timing shifts (change per decade in day-of-year (DOY) to half annual flow), where decreasing trends suggest more annual flow is occurring earlier in the year (a), and the percent change per decade in minimum (b), maximum (c) and median (d) annual flow. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
\label{fig:3}
\end{figure}

Our null model approach illustrates that support for a river network portfolio effect is greater than 90\% for all four annual flow trend metrics and as high as 98\% for median-annual-flow trends (Figure~\ref{fig:3}). Trend attenuation was; 3.1 (0.6--9.8 this and hereafter are 90\% intervals), 4.7 (1.1--14.7), 3.2 (0.9--9.0) and 5.2 (1.4--14.4) times greater than the null model for annual-flow timing, and minimum, maximum and median annual-flow, respectively. These statistics provide strong support for the hypothesis that river network portfolio effects contribute to climate dampening in large rivers.

@@

Regional flow-trends and attenuation strength varied by season. Analysis of monthly trends revealed that winter flows are getting higher and summer flows are getting lower over time, but that these trends have been more variable in smaller catchments (Figure~\ref{fig:4}, Figures~S6, S7). For example, winter maximum flows in small catchments ranged from nearly no change to dramatic 47\% increases per decade with decreases only as large as 9\%. Despite these extreme locations, on average the basin has only experienced moderate winter maximum flow increases of 5-9\%. Summer flows exhibited the opposite, with decreases in high flows over time. However, there was weak evidence of attenuation in the spring (Figure~\ref{fig:4}, Figures.~S6, S7). Overall, climate portfolios tended to have greater attenuation during stable periods of the year and decreased during seasonal transitions. In the spring, the likelihood of network-driven attenuation of maximum flow trends was as low as 24\% (Figure~\ref{fig:4}). The deterioration of the network's attenuation strength was coincident with the spring freshet, when snowmelt drives high flows across the basin. Synchronization events such as the freshet may subvert river network dampening mechanisms by homogenizing the region's response to seasonally driven climate shifts.

\begin{figure}[h]
\centering
\includegraphics[width=30pc]{Fig5_MaxMonth.pdf}
	\caption{Monthly maximum flow trend attenuation within the Fraser River basin. (\textbf{Left}) Fraser River's basin-wide maximum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulations. Observed values falling further from the center of the density distribution suggest greater evidence for changes in maximum flow and a greater shift in magnitude. (\textbf{Center}) Observed monthly Fraser River maximum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates. Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed. (\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months. These reflect months in the prior columns and describe the variation in percent change per decade of maximum flow among sites. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
\label{fig:4}
\end{figure}

@@

The Fraser River basin has seen changes in landcover between 1970 and 2007, which have impacted river discharge \citep[e.g.,][]{Zhang:2014}. Over this period, timber harvest has been the dominant contributor to changes in forest cover but increasingly natural events are having a greater impact (Figure~S8). For instance, abnormally hot, dry weather in 2003 led to fires that destroyed an unprecedented 2,600 km\textsuperscript{2} of forest largely within the Fraser River basin \citep{Filmon:2003}. Beginning in the early 2000's, western mountain pine beetle outbreaks began consuming large swaths of forest, impacting 20\% of British Columbia by 2013 \citep{schnorbus:2010}. These natural events are facilitated by climate change and are becoming increasingly common and extreme \citep{Melillo:2014, Maness:2013}. Pine beetle outbreaks are often subsequently heavily logged and timber removal can occur in anticipation of fires, resulting in harvest practices responding dynamically to climate change. Analyses indicated that while timber harvest rates did have some local impacts on hydrological change, harvest did not have substantial impacts on flows in a global model (Figure S3); thus we focus on long-term hydrological change and make the simplifying assumption that different forest harvest rates are subsumed by long-term climate change. Indeed, timber harvest was relatively consistent over the last 38 years across the vast majority of our 55 study catchments, with the exception of four relatively small, heavily fire and pine beetle impacted sites (Figure S8). Therefore, our study captures the interwoven direct and indirect impacts of climate change by considering changes in precipitation but also latent affects through climate filtering of altered landscapes due to climate change.

@@

Climate change is causing economic and conservation challenges for river systems worldwide \citep{Palmer:2009,Pecl:2017}. In the Fraser River, sockeye salmon have seen as much as 80\% pre-spawn mortality in years with later run-off and elevated temperatures, leading managers to reduce or close the fishery (e.g., 2013 and 2015). In the absence of climate dampening, the impacts of these climate extremes could have been considerably worse. In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage) or physically absorb climate change impacts (e.g., mangroves), river network climate dampening smooths out extreme climatic trends thereby reducing the impact of local extremes. But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \citep{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends. For example, dams synchronize the flow regimes of rivers \citep{Poff:2007} and logging increases the magnitude and frequency of extreme events such as high and low flows through increased runoff rates \citep{Zhang:2014}. These anthropogenic activities may magnify the impacts of climatic shifts such as the transition from snow- to rain-dominated precipitation in some basins. The climate cost of these watershed activities should be considered in environmental decision making. Climate change is a global challenge; here we suggest large rivers dampen local change by leveraging landscape diversity into a portfolio of climate, an important tool in the climate-mitigation toolbox.

@@

%  ACKNOWLEDGMENTS
\acknowledgments
This paper was possible only through the extensive use of many open source, freely available tools. All flow data analysis was done using R v3.3.2 (\url{www.r-project.org/}) and all R code can be found at \url{http://github.com/kchezik/River-Network-Flow-Trends.git}. Spatial analysis of watersheds was done using GDAL/OGR v1.11.5, and Whitebox GAT v3.4 \citep{Lindsay:2016} with visualization help from QGIS v2.18.2 (\url{www.qgis.org/}). Climate data were gathered using ClimateWNA v5.4 \citep{Wang:2016} and summary statitics were calculated using R and the \textit{rasterstats} module (\url{https://github.com/perrygeo/python-rasterstats.git}) in Python v2.7. Data processing was facilitated greatly by GNU Parallel \citep{Tange2011a}. This study could not have been done without the decades of flow data collected by those at Environment Canada. Raw HYDAT data can be obtained via Environment Canada or a simplified version can be found on GitHub. K.A. Chezik and J.W. Moore were supported by the Liber Ero Chair of Coastal Science and Management and Simon Fraser University. S.C. Anderson was supported by the David H. Smith Conservation Research Fellowship.

>>>>

%Packages
\usepackage{upgreek}
\usepackage{gensymb}
\usepackage{color,soul}

@@

\begin{abstract}
	River networks may dampen local hydrologic signals of climate change through the aggregation of upstream climate portfolio assets. Here we examine this hypothesis using flow and climate trend estimates (1970--2007) at 55 hydrometric gauge stations and across their contributing watersheds' within the Fraser River basin in British Columbia, Canada. Using a null hypothesis framework, we compared our observed attenuation of river flow trends as a function of increasing area and climate trend diversity, with null-simulated estimates to gauge the likelihood and strength of our observations. We found the Fraser River reduced variability in downstream long-term discharge by $>$91\%, with $>$3.1 times the attenuation than would be expected under null-simulation. Although the strength of dampening varied seasonally, our findings indicate that large free-flowing rivers offer a powerful and largely unappreciated process of climate change mitigation. River networks that integrate a diverse climate portfolio can dampen local extremes and offer climate change relief to riverine biota.
\end{abstract}

@@

\begin{figure}[h]
\centering
\includegraphics[width=35pc]{Fig1_Map.pdf}
	\caption{Climate trends (1970 - 2007) in mean annual temperature (MAT) and mean annual precipitation (MAP) within the Fraser River basin overlaid on a digital elevation model of British Columbia, Canada. Flow gauge sites (dots) are scaled by the size of the contributing area. The map is projected in Albers BC UTMs to provide an equal area depiction of the region but labels are expressed in WGS84 latitude and longitude.}
\label{fig:1}
\end{figure}

Here we consider whether river networks attenuate local long-term climate change through analysis of hydrometric data from one of the largest free-flowing rivers in North America. Located in British Columbia, Canada, the Fraser River drains an area approximately the size of the United Kingdom ($\sim$217,000 km\textsuperscript{2}) and in a rare combination is fairly well-monitored while also having no dams on its main stem \citep{Vorosmarty:2010} (Figure~\ref{fig:1}). This watershed drains interior high plateau, coastal mountains, and the Canadian Rockies, and thus integrates a diverse mosaic of landscapes, weather, and climate. As with other mid-latitude rivers \citep{Bindoff:2013}, the Fraser River's discharge volume and variability have increased over the last several decades \citep{Dery:2012,Morrison:2002}. Following projected patterns of climate change, the Fraser River basin is increasingly shifting from a snow dominated to rain-snow hybrid system \citep{Kang:2016}. Using 38 years of flow data collected concurrently at 55 hydrometric gauge stations throughout the watershed, we use a novel analytical approach to examine whether this large river exhibits signals of long-term climate dampening.

@@

	To assess signals of climate change in hydrology within the Fraser River basin we downloaded daily flow data from Environment Canada's HYDAT database. Within HYDAT we selected Fraser River basin hydro-metric gauge stations that were not observed to be dam influenced and which collected data in each month between 1970 and 2007; where no month was missing more than 5 days data and at least 35 complete years were present. We selected this 38-year range because this period maximized the number of gauge stations in the Fraser River basin (\textit{n}=55) operating concurrently over a long enough time period in which climate trends may be observable. We linearly interpolated missing data in log-space, but of the 746,544 days of data only 23 days were interpolated. These daily data were smoothed using a five-day rolling average before summarizing into annual and monthly response variables. This smoothing technique was used to reduce the influence of erroneous and unusually extreme values. Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and sites that were observed downstream of, and atypically close to, a dam (\textit{n}=3). The downstream distance of gauge stations from dams was determined using GIS spatial layers in the \textit{BC Data Catalog}, maintained by the BC provincial goverment.

	As climate change shifts precipitation patterns from snow to rain in temperate regions, rivers are generally predicted to have lower low-flows and higher high-flows, as well as earlier snowmelt \citep{Nijssen:2001}. These manifestations of rising temperatures are impacting species \citep{Xenopoulos:2006} and people \citep{Hirabayashi:2013}. We summarized our daily rolling average discharge data into annual and monthly flow metrics that capture these transitions. We used maximum- and minimum-flow estimates to capture trends in extreme discharge and median-flow estimates to capture general annual and seasonal trends. Because timing of flow is particularly important to the phenology of many organisms, we also calculated the day-of-year in which half the annual flow was reached.

@@

	Spatial variation in climate change would suggest catchments of different size integrate different amounts of climate variability. In order to calculate a climate index for each gauge station we delineated flow-gauge catchments and summarized climate trends within those catchments. Using Whitebox Geospatial Analysis Tools \citep{Lindsay:2016}, we delineated the Fraser River basin and the catchment areas of each flow-gauge station. Digital elevation models were provided by the provincial government of British Columbia, Canada and state government of Washington, USA and were pre-processed with a breaching algorithm and stream burned to facilitate proper flow path and accumulation models \citep{Woodrow:2016}. We summarized the climate variability integrated at each flow gauge by calculating the standard deviation in climate trends within their catchments using the \textit{rasterstats} module (\url{https://github.com/perrygeo/python-rasterstats.git}) in Python v2.7. The standard deviations of these climate variables were scaled between 0 and 1 and summed as a general climate index for each catchment.

@@

We hypothesize that flow gauge sites with larger climate portfolios will exhibit greater long-term flow trend stability than sites with smaller climate portfolios. We used catchment area as a proxy for the climate portfolio diversity of a given site, given that larger catchments have more variable climate portfolios (Figure~\ref{fig:2}) as well as aggregate more water and carry more weight on downstream dampening. Therefore, to test whether networks dampen climate signals, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
\begin{linenomath*}
\begin{equation}
	\hat{b}_{s} = c + d\sqrt{\mathrm{A}_{s}} + \eta_{s}, \quad
  \eta_{s} \sim \mathcal{N}(0, f(\mathrm{A}_{s})) \label{eq3},
\end{equation}
\end{linenomath*}
where $\hat{b}_{s}$ represents a flow trend at a given site ($s$) and $\mathrm{A}_{s}$ represents site $s$'s watershed area. We adjusted $\sqrt{\mathrm{A}_{s}}$ by subtracting the mean $\sqrt{\mathrm{A}}$, thereby centering our predictor. Fitted $d$ and $c$ parameters represent the mean effect of watershed area ($\sqrt{\mathrm{A}}$), on flow trends and the mean flow trend for a $\sim$5,700 km\textsuperscript{2} watershed, respectively. Captured in our climate portfolio proxy variable ($\mathrm{A}$), is the interaction between climate and landuse. Timber harvest, the dominant source of land cover change in the Fraser River basin, scales with sub-basin area and contributes to changes in flow. Although exploratory linear models at the site level do suggest an effect of timber harvest on long-term flow trends for some gauge stations (Figure~S1, S2), when information is shared across sites in a global model, climate effects dominate (Figure~S3). Thus we did not attempt to separate the contributions of land cover change, climate change and their interaction on long-term flow trends.

@@

Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 92 and 96\% dampening of flow trends moving from headwater catchments to the network's outlet. The largest catchments were approximately 10 times more stable in their climate response than the smallest catchments. For instance, trends ranged between a 6\% reduction and 19\% increase in median flow per decade among small watersheds, while large watersheds ranged between 0.8 and 1.6\% change per decade. Similar attenuation was seen in flow timing with a 94\% reduction in the day-of-year to half-annual-flow trends such that small watersheds were reaching half their annual flow between 8 days later and 25 days earlier while large watersheds have only shifted 2 to 4 days earlier between 1970 and 2007.

\begin{figure}[h]
\centering
\includegraphics[width=35pc]{Fig4_AnnFun.pdf}
	\caption{Annual flow trend attenuation within the Fraser River basin. (\textbf{Left}) Trend estimates ($\hat{b}_{s}$) $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large). Blue lines represent observed attenuation; orange and red represent simulated attenuation that is weaker and stronger than observed. (\textbf{Right}) Density plots show null model simulated variance exponents ($\hat{\updelta}$) and the proportion on either side of the observed variance exponent (blue). Flow metrics include long-term flow-timing shifts (change per decade in day-of-year (DOY) to half annual flow), where decreasing trends suggest more annual flow is occurring earlier in the year (a), and the percent change per decade in minimum (b), maximum (c) and median (d) annual flow. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
\label{fig:3}
\end{figure}

Our null model approach illustrates that support for a river network portfolio effect is greater than 90\% for all four annual flow trend metrics and as high as 98\% for median-annual-flow trends (Figure~\ref{fig:3}). Trend attenuation was 3.1 (0.6--9.8 this and hereafter are 90\% intervals), 4.7 (1.1--14.7), 3.2 (0.9--9.0) and 5.2 (1.4--14.4) times greater than the null model for annual-flow timing, and minimum, maximum and median annual-flow, respectively. These statistics provide strong support for the hypothesis that river network portfolio effects contribute to climate dampening in large rivers.

@@

Regional flow-trends and attenuation strength varied by season. Analysis of monthly trends revealed that winter flows are getting higher and summer flows are getting lower over time, but that these trends have been more variable in smaller catchments (Figure~\ref{fig:4}, Figures~S6, S7). For example, winter maximum flows in small catchments ranged from nearly no change to dramatic 47\% increases per decade with decreases only as large as 9\%. Despite these extreme locations, on average the basin has only experienced moderate winter maximum flow increases of 5-9\%. Summer flows exhibited the opposite, with decreases in high flows over time. However, there was weak evidence of attenuation in the spring (Figure~\ref{fig:4}, Figures.~S6, S7). Overall, climate portfolios tended to have greater attenuation during stable periods of the year and decreased during seasonal transitions. In spring, the likelihood of network-driven attenuation of maximum flow trends was as low as 24\% (Figure~\ref{fig:4}). Deterioration of the network's attenuation strength was coincident with the spring freshet, when snowmelt drives high flows across the basin. Synchronization events such as the freshet may subvert river network dampening mechanisms by homogenizing the region's response to seasonally driven climate shifts.

\begin{figure}[h]
\centering
\includegraphics[width=30pc]{Fig5_MaxMonth.pdf}
	\caption{Monthly maximum flow trend attenuation within the Fraser River basin. (\textbf{Left}) Fraser River's basin-wide maximum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulations. Observed values falling further from the center of the density distribution suggest greater evidence for changes in maximum flow and a greater shift in magnitude. (\textbf{Center}) Observed monthly Fraser River maximum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates. Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed. (\textbf{Right}) Trend estimates ($\hat{b}_{s}$) $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months. These reflect months in the prior columns and describe the variation in percent change per decade of maximum flow among sites. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
\label{fig:4}
\end{figure}

@@

The Fraser River basin has seen changes in landcover between 1970 and 2007, which have impacted river discharge \citep[e.g.,][]{Zhang:2014}. Over this period, timber harvest has been the dominant contributor to changes in forest cover (Figure~S8) but increasingly natural events are having a greater impact. For instance, abnormally hot, dry weather in 2003 led to fires that destroyed an unprecedented 2,600 km\textsuperscript{2} of forest largely within the Fraser River basin \citep{Filmon:2003}. Beginning in the early 2000's, western mountain pine beetle outbreaks began consuming large swaths of forest, impacting 20\% of British Columbia by 2013 \citep{schnorbus:2010}. These natural events are facilitated by climate change and are becoming increasingly common and extreme \citep{Melillo:2014, Maness:2013}. Pine beetle outbreaks are often subsequently heavily logged and timber removal can occur in anticipation of fires, resulting in harvest practices responding dynamically to climate change. Analyses indicated that while timber harvest rates did have some local impacts on hydrological change, harvest did not have substantial impacts on flows in a global model (Figure S3); thus we focus on long-term hydrological change and make the simplifying assumption that different forest harvest rates are subsumed by long-term climate change. Indeed, timber harvest was relatively consistent over the last 38 years across the vast majority of our 55 study catchments, with the exception of four relatively small, heavily fire and pine beetle impacted sites (Figure S8). Therefore, our study captures the interwoven direct and indirect impacts of climate change by considering changes in precipitation but also latent affects through climate filtering of altered landscapes due to climate change.

@@

Climate change is causing economic and conservation challenges for river systems worldwide \citep{Palmer:2009,Pecl:2017}. In the Fraser River, sockeye salmon have seen as much as 80\% pre-spawn mortality in years with later run-off and elevated temperatures, leading managers to reduce or close the fishery (e.g., 2013 and 2015). In the absence of climate dampening, the impacts of these climate extremes could have been considerably worse. In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage, \citep{Jones:2012}) or physically absorb climate change impacts (e.g., mangroves, \citep{Arkema:2013}), river network climate dampening smooths out extreme climatic trends thereby reducing the impact of local extremes. But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \citep{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends. For example, dams synchronize the flow regimes of rivers \citep{Poff:2007}. These anthropogenic activities may magnify the impacts of climatic shifts such as the transition from snow- to rain-dominated precipitation in some basins. The climate cost of these watershed activities should be considered in environmental decision making. Climate change is a global challenge; here we suggest large rivers dampen local change by leveraging landscape diversity into a portfolio of climate, an important tool in the climate-mitigation toolbox.

@@

%  ACKNOWLEDGMENTS
\acknowledgments
This paper relied on many open source tools including R v3.3.2 (\url{www.r-project.org/}), Python v2.7, QGIS v2.18 and WhiteboxGAT \citep{Lindsay:2016}. All R code can be found at \url{http://github.com/kchezik/River-Network-Flow-Trends.git}. Data processing was facilitated greatly by GNU Parallel \citep{Tange2011a}. This study relied on decades of flow data collected by those at Environment Canada. Raw HYDAT data can be obtained via Environment Canada or a simplified version can be found on GitHub. K.A. Chezik and J.W. Moore were supported by the Liber Ero Chair of Coastal Science and Management and Simon Fraser University. S.C. Anderson was supported by the David H. Smith Conservation Research Fellowship.
